---
title: "GroupProject"
output: html_document
date: "2023-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dbscan)
library(tidyverse)
library(geosphere)
library(ggplot2)
library(lubridate)
library(ggmap)
library(mapproj)
library(leaflet)
library(sp)
```



```{r}
movements0 = read.csv("CleanData/movements_0.csv", sep=",", header = TRUE)

```


# removing moving points

```{r}
timeWindow = 15

prefilteredMovements0 <- movements0 %>%
  mutate(forwardDistance = 
           ifelse(!is.na(lead(date, n=timeWindow)), distHaversine(
             cbind(longitude, latitude),
             cbind(lead(longitude, n=timeWindow), lead(latitude, n = timeWindow))
           ),
           NA_real_
           )) %>%
  mutate(backwardDistance = 
           ifelse(!is.na(lag(date, n=timeWindow)), distHaversine(
             cbind(longitude, latitude),
             cbind(lag(longitude, n=timeWindow), lag(latitude, n = timeWindow))
           ),
           NA_real_
           ))
ggplot(prefilteredMovements0, aes(x=forwardDistance)) +
  geom_histogram(binwidth = 25)
  
filteredMovements0 <- prefilteredMovements0 %>% filter(forwardDistance < 75 | backwardDistance < 75)
```

# print a graph 
```{r}
#print("hello")
range(filteredMovements0$forwardDistance, na.rm = TRUE)


ggplot(filteredMovements0, aes(x=forwardDistance)) +
  geom_histogram(binwidth = 25)
```
# Resampling data to 5 minute intervals (this takes a while)

```{r}
newFiltered0 = data.frame(date=character(), time=character(), longitude=double(), latitude=double(), samples=double())
i = 1
print(filteredMovements0[1,5])
while (i <= length(filteredMovements0[,1])){
  minutes = as.numeric(format(strptime(filteredMovements0[i,5], "%H:%M:%S"), '%M'))
  date = filteredMovements0[i,4]
  time = filteredMovements0[i,5]
  
  sumLat = filteredMovements0[i,3]
  sumLong = filteredMovements0[i,2]
  
  i = i + 1
  num = 1
  curMin =  as.numeric(format(strptime(filteredMovements0[i,5], "%H:%M:%S"), '%M'))
  while ((curMin - minutes) < 5 & num < 10 & i <= length(filteredMovements0[,1])) {
    sumLat = sumLat + filteredMovements0[i,3]
    sumLong = sumLong + filteredMovements0[i,2]
    i = i + 1
    num = num + 1
    curMin =  as.integer(format(strptime(filteredMovements0[i,5], "%H:%M:%S"), '%M'))
  }
  newLat = sumLat / num
  newLong = sumLong / num
  
  newFiltered0 <- rbind(newFiltered0, list(date, time, newLong, newLat, num))
}

colnames(newFiltered0) <- c('date', 'time', 'longitude', 'latitude', 'samples')
```
# create coordinate0 (no longer needed?)

```{r}
coordinates0 <- filteredMovements0 %>%
  select(longitude, latitude) %>%
  filter(row_number() %% 5 == 0)
c0 <- hdbscan(coordinates0, minPts = 5)
```

# perform hdbscan on resampled data

```{r}
coordinates0 <- newFiltered0 %>%
  select(longitude, latitude)
c0 <- hdbscan(coordinates0, minPts = 20)
```

# perform dbscan on resampled data
eps is the radius that we search for neighboring points within
minpts is the minimum number of points to form a cluster

```{r}
dc0 <- dbscan(coordinates0, eps = 0.001, minPts = 5)
basicCoordinates <- coordinates0
basicCoordinates$clusterLabel <- as.factor(dc0$cluster)
```


# add labels to data

```{r}
coordinates0$clusterLabel <- as.factor(c0$cluster)
```

# create 1 point for each cluster
```{r}
testCluster <- subset(coordinates0, clusterLabel == 0)
```

```{r}
clusterCoordinates = data.frame(longitude=double(), latitude=double(), count=double())

length = length(unique(coordinates0$clusterLabel))
i = 0

while (i < length) {
  tempCluster = coordinates0 [coordinates0$clusterLabel == i,]
  
  tempLat = mean(tempCluster$latitude)
  tempLong = mean(tempCluster$longitude)
  num = nrow(tempCluster)
  
  i = i + 1
  
  clusterCoordinates = rbind(clusterCoordinates, 
                             list(tempLong, tempLat, num))
}

colnames(clusterCoordinates) <- c('longitude', 'latitude', 'samples')
```

# plot via leaflet

```{r}
newClusterCoordinates <- clusterCoordinates
coordinates(newClusterCoordinates) <- ~longitude+latitude
```
```{r}
leaflet(newClusterCoordinates) %>% addMarkers() %>% addTiles()
```

for (i in 1:nrow(clusterCoordinates)) {
  map <- addCircleMarkers(map, 
                          lng = clusterCoordinates$longitude[i],
                          lat = clusterCoordinates$latitude[i],
                          radius = clusterCoordinates$samples[i]/1000.0+1,
                          color = "blue")
}
```{r}

map <- leaflet() %>%
  setView(lng = mean(clusterCoordinates$longitude), lat = mean(clusterCoordinates$latitude), zoom = 11) %>%
  addTiles()

for (i in 1:nrow(clusterCoordinates)) {
  map <- addCircleMarkers(map, 
                          lng = clusterCoordinates$longitude[i],
                          lat = clusterCoordinates$latitude[i],
                          radius = clusterCoordinates$samples[i]^(1/3)/1.5,
                          color = "blue")
}

map
```
# ---------------------------------------------
# Simplify dbscan data to one point per cluster
```{r}
basicClusterCoordinates = data.frame(longitude=double(), latitude=double(), count=double())

length = length(unique(basicCoordinates$clusterLabel))
i = 0

while (i < length) {
  tempCluster = basicCoordinates [basicCoordinates$clusterLabel == i,]
  
  tempLat = mean(tempCluster$latitude)
  tempLong = mean(tempCluster$longitude)
  num = nrow(tempCluster)
  
  i = i + 1
  
  basicClusterCoordinates = rbind(basicClusterCoordinates, 
                             list(tempLong, tempLat, num))
}

colnames(basicClusterCoordinates) <- c('longitude', 'latitude', 'samples')
```
# plot clusters from DBSCAN

```{r}

map <- leaflet() %>%
  setView(lng = mean(basicClusterCoordinates$longitude), lat = mean(basicClusterCoordinates$latitude), zoom = 11) %>%
  addTiles()

for (i in 1:nrow(basicClusterCoordinates)) {
  map <- addCircleMarkers(map, 
                          lng = basicClusterCoordinates$longitude[i],
                          lat = basicClusterCoordinates$latitude[i],
                          radius = basicClusterCoordinates$samples[i]^(1/3)/1.5,
                          color = "blue")
}

map
```




# stamen maps no longer exist, google requires API
```{r}
map <- get_stadiamap(bbox = c(lon = range(coordinates0$longitude), 
                              lat = range(coordinates0$latitude)),
                              zoom = 10,
                              maptype = "toner")
ggmap(map) + 
  geom_point(data = coordinates0, aes(x = longitude, y = latitude, color = as.factor(cluster_label)), size = 2) +
  scale_color_discrete(name = "Cluster") +
  labs(title = "Location Clustering Map")
```



```{r}
map("world", fill = TRUE, col = "white", bg = "lightblue", xlim = c(-122.5, -122.2), ylim = c(47.5, 47.7), resolution = 0)

#This functions, but all points plotted as same color
#points(coordinates0$longitude, coordinates0$latitude, col = "black", bg = "red", pch = 21)

#makes set of colors for each cluster
num_clusters <- length(unique(coordinates0$clusterLabel))
colors <- rainbow(num_clusters)

points(coordinates0$longitude, coordinates0$latitude, col = "black", bg = colors[coordinates0$clusterLabel], pch = 21)

#plots all points
#for (cluster_id in unique(coordinates0$clusterLabel)) {
#  cluster_data <- subset(coordinates0, clusterLabel == cluster_id)
#  points(cluster_data$longitude, cluster_data$latitude, col = colors[cluster_id], bg = #colors[cluster_id], pch = 21)
#}

#for(cluster_id in unique(coordinates0$clusterLabel)) {
#  clusterData <- subset(coordinates0, clusterLabel == cluster_id)
#  points(clusterData$longitude, clusterData$latitude, col = #rainbow(length(unique(coordinates0$clusterLabel)))[cluster_id], pch = #20)
#}
```



NOT WORKING...

```{r}
#map("world", fill = TRUE, col = "white", bg = "lightblue", xlim = range(coordinates0$longitude), ylim = range(coordinates0$latitude))


#for(cluster_id in unique(coordinates0$clusterLabel)) {
#  clusterData <- subset(coordinates0, clusterLabel == cluster_id)
#  points(clusterData$longitude, clusterData$latitude, col = #rainbow(length(unique(coordinates0$clusterLabel)))[cluster_id], pch = 20)
#}

world <- map_data("world")

map_plot <- ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "lightblue", color = "white") +
  theme_void()

points_plot <- map_plot +
  geom_point(data = coordinates0, aes(x = longitude, y = latitude, color = clusterLabel), size = 2) +
  scale_color_discrete(name = "Cluster") +
  labs(title = "HDBSCAN Clustering with mapproj and ggplot2")

print(points_plot)
```

```{r}
write.csv(newFiltered0, "CleanData/FilteredData0.csv", row.names = FALSE)
```


```{r}
topClusters <- head(clusterCoordinates[order(-clusterCoordinates$samples), ], 30)

topClusters <- topClusters %>% mutate(rank = row_number())



ggplot(topClusters, aes(x = rank)) + 
  geom_bar() +
  labs(title = "Most Common Stay Points", x = "Stay Points", y = "Count of Measurements")
```

```{r}
write.csv(topClusters, "TopClusters.csv", row.names = FALSE)
```

```{r}
write.csv(clusterCoordinates, "clusterCoordinates.csv", row.names = FALSE)
```

```{r}
basicTopClusters <- head(basicClusterCoordinates[order(-basicClusterCoordinates$samples), ], 30)

basicTopClusters <- basicTopClusters %>% mutate(rank = row_number())
write.csv(basicTopClusters, "basicClusterCoordinates.csv", row.names = FALSE)
```



